---
title: "Piracy Notebook"
author: "Gavin McDonald"
output: 
  html_notebook: 
    number_sections: yes
    toc: yes
date: "January 18, 2017"
---
# Introduction

# Methods
## Load necessary libraries
```{r load_libraries}
library(rgdal)
library(sp)
library(rgeos)
library(mapproj)
library(leaflet)
library(googlesheets)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(readr)
library(broom)
library(geosphere)
library(raster)
library(bigrquery)
```
## Load anti-shipping activity messages
Data from http://msi.nga.mil/NGAPortal/MSI.portal?_nfpb=true&_pageLabel=msi_portal_page_65
```{r load_data, results='hide',error=FALSE,message=FALSE,warning=FALSE}
## load anti-shipping data
ASAM <- readOGR(dsn = "data/ASAM_shp", layer = "ASAM 18 JAN 17", stringsAsFactors = FALSE)
## Process dates, and filter out anything before 2012
ASAM@data <- ASAM@data %>%
  mutate(Date = as_date(DateOfOcc),
         Year = year(Date)) %>%
  dplyr::select(-DateOfOcc)

vesselTypes <- unique(ASAM@data$Victim)[grepl("Fish",unique(ASAM@data$Victim)) | grepl("FISH",unique(ASAM@data$Victim)) | grepl("fish",unique(ASAM@data$Victim))]

ASAM <- ASAM %>%
  subset(Year > 2012)
  # subset(Victim %in% vesselTypes)
## Remove duplicates
ASAM <- remove.duplicates(ASAM)
## read in EEZ
eez <- readOGR(dsn = "data/World_EEZ_v9_20161021_LR", layer = "eez_lr", stringsAsFactors = FALSE)

## set projection
proj4string(ASAM)<-proj4string(eez)

## look only at events happening within eezs
eezASAMOverlap <- overlay<-raster::intersect(ASAM,eez)
## events in high seas
highSeasOverlap <- subset(ASAM,!(Reference %in% eezASAMOverlap@data$Reference))

highSeasOverlap@data <- highSeasOverlap@data %>%
  mutate(PolygonID = NA,
         MRGID = NA,
         GeoName = "High Seas",
         Pol_type = NA,
         MRGID_Ter1 = NA,
         Territory1 = NA,
         MRGID_Sov1 = NA,
         Sovereign1 = NA,
         ISO_Ter1 = NA,
         MRGID_Ter2 = NA,
         Territory2 = NA,
         MRGID_Sov2 = NA,
         Sovereign2 = NA,
         ISO_Ter2 = NA,
         MRGID_Ter3 = NA,
         Territory3 = NA,
         MRGID_Sov3 = NA,
         Sovereign3 = NA,
         ISO_Ter3 = NA,
         x_1 = NA,
         y_1 = NA,
         Area_km2 = NA,
         MRGID_EEZ = NA)
  
masterAttacks <- eezASAMOverlap %>%
  rbind(highSeasOverlap) %>%
  remove.duplicates()
## set world base map
#world <- map("world", fill = TRUE, plot = FALSE)
world <- readOGR(dsn = "data/TM_WORLD_BORDERS-0.3", layer = "TM_WORLD_BORDERS-0.3", stringsAsFactors = FALSE) %>%
  gUnaryUnion(id = .@data$country)
detach("package:rgdal", unload=TRUE)
```

## Plot attack data
```{r attackPlot}

data.frame(country=gsub(" Exclusive Economic Zone"," EEZ",masterAttacks@data$GeoName)) %>%
  group_by(country) %>%
  summarize(n = n()) %>%
  #filter(n > 10) %>%
  ggplot(aes(x=reorder(country,-n),y=n)) +
           geom_histogram(stat="identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Country") +
  ylab("Number of attacks\n2012 - 2016")

```


## Set up buffers and controls
```{r buffer}
# ## Just look at Somalia
# somaliPiracy <- masterAttacks %>%
#   #subset(Aggressor == "Pirates") %>%
#   subset(GeoName == "Somali Exclusive Economic Zone")
# 
# eezSomalia <- eez %>%
#   subset(GeoName == "Somali Exclusive Economic Zone")

impactZones <- gBuffer(masterAttacks,width=0.5,byid=TRUE, quadsegs=1, capStyle="SQUARE") %>%
  SpatialPolygonsDataFrame(data=masterAttacks@data,match.ID=TRUE)
impactZones@data$Control <- "Impact"

controlZones <- gBuffer(masterAttacks,width=1,byid=TRUE, quadsegs=1, capStyle="SQUARE") %>%
  SpatialPolygonsDataFrame(data=masterAttacks@data,match.ID=TRUE)

controlZones<- controlZones - impactZones
controlZones@data$Control <- "Control"

## Only look inside Somalia EEZ
#impactZones <- raster::intersect(impactZones,eezSomalia)
#controlZones <- raster::intersect(controlZones,eezSomalia)

impactAreas <- lapply(impactZones@data$Reference,function(x){
  temp<-SpatialPolygons((subset(impactZones,Reference==x)@polygons)) %>%
      areaPolygon() %>%
      as_data_frame() %>%
      mutate(Reference=x,
             area_km2=value/1e6,
             Control="Impact")
}) %>%
  bind_rows()

controlAreas <- lapply(controlZones@data$Reference,function(x){
  temp<-SpatialPolygons((subset(controlZones,Reference==x)@polygons)) %>%
      areaPolygon() %>%
      as_data_frame() %>%
      mutate(Reference=x,
             area_km2=value/1e6,
             Control="Control")
}) %>%
  bind_rows()

masterAreas <- bind_rows(impactAreas,controlAreas) %>%
  left_join(masterAttacks@data %>%
              dplyr::select(Reference,AttackDate = Date))
```


## Plot locations of attacks

```{r plot}

leaflet(data = eez) %>%
  # fitBounds(0, 60, 35, 72) %>%
  addPolygons(
    fillOpacity = 0.15, 
    fillColor = "blue", 
    stroke=TRUE,
    weight=1,
    color="black",
    popup = ~paste0("<font color=#000000>",
      "<b>EEZ: </b>", GeoName, "<br>",
    "</font>"))  %>%
  addPolygons(data=impactZones,
              color="red",
              stroke=FALSE,
              fillOpacity = 0.5,
              popup = ~paste0("<font color=#000000>",
                "<b>Reference ID: </b>", Reference, "<br>",
                "<b>EEZ: </b>", GeoName, "<br>",
                "<b>Date: </b>", Date, "<br>",
                "<b>Aggressor: </b>", Aggressor, "<br>",
                "<b>Victim: </b>", Victim, "<br>",
                "<b>Description: </b>", gsub("NA","",paste(Desc1,Desc2,Desc3,Desc4,Desc5,Desc6,sep=" ")), "<br>",
                 "</font>")) %>%
  # addPolygons(data=controlZones,
  #             color="green",
  #             stroke=FALSE,
  #             fillOpacity = 0.5) %>%
  addPolygons(
    data = world,
    group = "world",
    fillColor="white",
    color="black",
    stroke=TRUE,
    weight=1,
    fillOpacity = 1)
```

## Load GFW data

```{r GFWdata}
## Create bounding box around each zone
# boxes <- lapply(controlZones@data$Reference,function(x){
#   temp <- subset(controlZones,Reference == x)
boxes <- lapply(impactZones@data$Reference,function(x){
  temp <- subset(impactZones,Reference == x)
  box <- bbox(temp) %>%
    as_data_frame
  
  box <- cbind(box[1,],box[2,])
  colnames(box) <- c("min_lon","max_lon", "min_lat","max_lat")
  box <- box %>%
    as_data_frame %>%
    mutate(Reference = x)
}) %>%
  bind_rows()

#GFW_connection <- src_bigquery("ucsb-gfw", "all_fishing_effort", max_pages = Inf)
#fao_file <- "all_fishing_effort_all_years"
grid_resolution_degree <- 0.25 # Set resolution of raster grid, degrees
#fishingEffort <- lapply(c("2016-209"), function(x){
fishingEffort <- lapply(masterAttacks@data$Reference, function(x){
  print(x)
  tempBox <- boxes[which(boxes$Reference == x),]
  max_lat <- tempBox$max_lat
  min_lat <- tempBox$min_lat
  max_lon <- tempBox$max_lon
  min_lon <- tempBox$min_lon
    sql <-  paste("SELECT
DATE(timestamp) Date,
SUM(hours) total_hours,
FROM
 all_fishing_effort.all_fishing_effort_all_years
WHERE
 measure_new_score >=.5
AND
hours > 0
 AND 
lat <",max_lat,"
AND
lat >",min_lat,"
AND
lon <",max_lon,"
AND
lon >",min_lon,"
AND
NOT (distance_from_shore < 1000 AND implied_speed < 1)
GROUP BY
Date",
sep="")
    project <-  "ucsb-gfw"
    binned_effort <- query_exec(sql,project,max_pages = Inf) %>%
      mutate(Reference = tempBox$Reference)
#   sql <-  paste("SELECT
#  FLOOR(lat/",grid_resolution_degree,") * ",grid_resolution_degree," lat_bin,
#  FLOOR(lon/",grid_resolution_degree,") * ",grid_resolution_degree," lon_bin,
# DATE(timestamp) Date,
# hours
# FROM
#  all_fishing_effort.all_fishing_effort_all_years
# WHERE
#  measure_new_score >=.5
# AND
# hours > 0
#  AND 
# lat <",max_lat,"
# AND
# lat >",min_lat,"
# AND
# lon <",max_lon,"
# AND
# lon >",min_lon,"
# AND
# NOT (distance_from_shore < 1000 AND implied_speed < 1)
# GROUP BY
# Date,
#  lat_bin,
#  lon_bin,
# hours",
# sep="")
#     project <-  "ucsb-gfw"
#     binned_effort <- query_exec(sql,project,max_pages = Inf) %>%
#       mutate(Reference = tempBox$Reference)
  # tbl(GFW_connection, from = fao_file) %>%
  #   dplyr::select(mmsi, timestamp, lon, lat, hours, measure_new_score,flag_country_name,eez_name,distance_from_shore,implied_speed) %>%
  #   filter(
  #     lat < max_lat &
  #       lat > min_lat &
  #       lon < max_lon &
  #       lon > min_lon &
  #       measure_new_score > 0.5 &
  #       !(distance_from_shore < 1000 & implied_speed < 1)
  #   ) %>%
  #   collect(n=Inf) %>%
  #   mutate(Reference = tempBox$Reference)
})   %>%
  ## bind over all attacks
  bind_rows()
save(fishingEffort,file="data/fishingEffort.Rdata")
load("data/fishingEffort.Rdata")
#coordinates(fishingEffort) <- c("lon_bin", "lat_bin")
#proj4string(fishingEffort) <- sp::proj4string(masterAttacks)

# daily_fishing_effort <- lapply(masterAttacks@data$Reference,function(x){
#   if(!(x %in% unique(fishingEffort$Reference))) return()
#   controlEffort <- fishingEffort[!is.na(over(fishingEffort,as(controlZones,"SpatialPolygons"))),]
#   impactEffort <- fishingEffort[!is.na(over(fishingEffort,as(impactZones,"SpatialPolygons"))),]
#   
#   controlEffort <- controlEffort@data %>%
#     mutate(Control = "Control")
#   
#   impactEffort <- impactEffort@data %>%
#     mutate(Control = "Impact")
#   
#   all_effort <- bind_rows(controlEffort,impactEffort)
# }) %>%
#   bind_rows() %>%
#   #mutate(Date = floor_date(timestamp, unit = "days")) %>%
#   group_by(Reference,Control,Date) %>%
#   summarize(Total_Hours = sum(hours)) %>%
#   left_join(masterAreas,by=c("Reference","Control")) %>%
#   mutate(hours_per_km2 = Total_Hours/area_km2)

daily_fishing_effort <- fishingEffort %>%
  mutate(Control = "Impact") %>%
  left_join(masterAreas,by=c("Reference","Control")) %>%
  rename(Total_Hours = total_hours) %>%
  mutate(hours_per_km2 = Total_Hours/area_km2)

all_dates = seq(as.Date("2012-01-01"), as.Date("2016-12-31"), by="day")

completeDF <- expand.grid(all_dates,c("Control","Impact"),masterAttacks@data$Reference)
colnames(completeDF) <- c("Date","Control","Reference")
daily_fishing_effort <- completeDF %>%
  as_data_frame() %>%
  mutate(Date = floor_date(Date, unit = "days")) %>%
  left_join(masterAreas,by=c("Reference","Control")) %>%
  left_join(daily_fishing_effort %>%
              mutate(Date = as_date(Date)) %>%
              dplyr::select(Date,Control,Reference,Total_Hours,hours_per_km2),by=c("Date","Control","Reference")) %>%
  mutate(Total_Hours = ifelse(is.na(Total_Hours),0,Total_Hours),
         hours_per_km2 = ifelse(is.na(hours_per_km2),0,hours_per_km2))

daily_fishing_effort <- daily_fishing_effort %>%
  left_join(masterAttacks@data %>%
              dplyr::select(Reference,
                     GeoName))
```

## Statistics
Here we perform some statistics on the daily fishing effort data. These include:  
  
* Use a 30-day rolling mean for daily fishing effort [hours/km^2]
  
* Get predicted values from a regression that includes calendar effects and ENSO effect
  
* Detrend the data by subtracting this predicted regression from the observed values   
  
```{r stats,error=FALSE,message=FALSE,warning=FALSE}
## read in enso data
enso <- 
  read_table("http://www.esrl.noaa.gov/psd/gcos_wgsp/Timeseries/Data/nino34.long.anom.data",
             na = c("-99.99", "99.99"), skip = 1, n_max = year(Sys.time()) - 1870 + 1,
             col_names = c("year", 1:12)) %>%
  filter(year >= min(year(daily_fishing_effort$Date))) %>%
  filter(year <= max(year(daily_fishing_effort$Date))) %>%
  gather(month, enso, -year) %>%
  mutate(month = as.double(month),
         enso = ifelse(is.na(enso),mean(enso,na.rm=TRUE),enso))

## add column for 30 day rolling mean, join on enso data
daily_fishing_effort <- daily_fishing_effort %>%
  mutate(fhours_30drm = zoo::rollmean(hours_per_km2, 30, fill=0),
         year = year(Date),
         month = month(Date)) %>%
  mutate(year = year(Date),
         month = month(Date)) %>%
  left_join(enso, by = c("year", "month")) %>% 
  mutate(zone = paste(Reference,Control,sep="_"))

seas_pred <-
  daily_fishing_effort %>%
  group_by(zone) %>%
  do(pred = predict(lm(fhours_30drm ~ factor(month) + factor(wday(Date)), data = .))) %>%
  ungroup()

daily_fishing_effort$pred <- unlist(seas_pred$pred)
daily_fishing_effort$fhours_dtrend <- daily_fishing_effort$hours_per_km2 - daily_fishing_effort$pred 
daily_fishing_effort$fhours_dtrend30drm = zoo::rollmean(daily_fishing_effort$fhours_dtrend, 30, fill=0)
```

```{r resultsDailyEffort, fig.width=8.5, fig.height=6,error=FALSE,message=FALSE,warning=FALSE}
# lapply(unique(fishingEffort$Reference),function(x){
#   
#   daily_fishing_effort_temp <- daily_fishing_effort %>%
#     subset(Reference == x) 
#   daily_fishing_effort_temp %>%
#     ggplot(aes(x=as_date(Date),y=fhours_30drm,col=Control,area=Control)) +
#     geom_line(lwd=0.8) +
#     geom_vline(aes(xintercept = as.numeric(as_date(AttackDate[1]))),linetype=2,lwd=1,col="blue") +
#   scale_x_date(limits = as.Date(c('2012-01-01','2016-12-31'))) +
#   xlab("") +
#   ylab("Fishing effort [hours/km^2]\nRolling 30-day mean") +
#   ggtitle(paste("Country: ",daily_fishing_effort_temp$GeoName[1],"\nAttack Reference ID: ",x,"\nDashed blue line represents attack date"))
# })

library(trelliscopejs)
cum_fishing_effort <- daily_fishing_effort %>%
  group_by(Reference) %>%
  summarize(cum_effort = sum(Total_Hours))

daily_fishing_effort %>%
  left_join(cum_fishing_effort,by=c("Reference")) %>%
  ggplot(aes(x=Date,y=Total_Hours)) +
  geom_line()  +
  xlab("") +
  ylab("Daily fishing effort [hours]") +
  geom_vline(xintercept = as.numeric(as_date(daily_fishing_effort$AttackDate[1])),linetype=2,lwd=1,col="darkorange") +
  facet_trelliscope(~ Reference, nrow = 3, ncol = 3,self_contained = TRUE,
                    state = list(sort = list(sort_spec("cum_effort",dir="desc"))))
```
