---
title: "Piracy Notebook"
author: "Gavin McDonald"
output: 
  html_notebook: 
    number_sections: yes
    toc: yes
date: "January 18, 2017"
---
# Introduction

# Methods
## Load necessary libraries
```{r load_libraries}
library(rgdal)
library(sp)
library(rgeos)
library(mapproj)
library(leaflet)
library(googlesheets)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(readr)
library(broom)
library(geosphere)
library(raster)
library(bigrquery)
```
## Load anti-shipping activity messages
```{r load_data, results='hide',error=FALSE,message=FALSE,warning=FALSE}
## load anti-shipping data
ASAM <- readOGR(dsn = "data/ASAM_shp", layer = "ASAM 18 JAN 17", stringsAsFactors = FALSE)
## Process dates, and filter out anything before 2012
ASAM@data <- ASAM@data %>%
  mutate(Date = as_date(DateOfOcc),
         Year = year(Date)) %>%
  dplyr::select(-DateOfOcc)

ASAM <- ASAM %>%
  subset(Year > 2012)
## Remove duplicates
ASAM <- remove.duplicates(ASAM)
## read in EEZ
eez <- readOGR(dsn = "data/World_EEZ_v9_20161021_LR", layer = "eez_lr", stringsAsFactors = FALSE)

## set projection
proj4string(ASAM)<-proj4string(eez)

## look only at events happening within eezs
eezASAMOverlap <- overlay<-raster::intersect(ASAM,eez)
## set world base map
world <- map("world", fill = TRUE, plot = FALSE)
```
## Set up buffers and controls
```{r buffer}
## Just look at Somalia
somaliPiracy <- eezASAMOverlap %>%
  #subset(Aggressor == "Pirates") %>%
  subset(GeoName == "Somali Exclusive Economic Zone")

eezSomalia <- eez %>%
  subset(GeoName == "Somali Exclusive Economic Zone")

impactZones <- gBuffer(somaliPiracy,width=0.5,byid=TRUE) %>%
  SpatialPolygonsDataFrame(data=somaliPiracy@data,match.ID=TRUE)
impactZones@data$Control <- "Impact"

controlZones <- impactZones %>%
  gBuffer(width=0.5,byid=TRUE) 

controlZones<- controlZones - impactZones
controlZones@data$Control <- "Control"

## Only look inside Somalia EEZ
impactZones <- raster::intersect(impactZones,eezSomalia)
controlZones <- raster::intersect(controlZones,eezSomalia)

impactAreas <- lapply(impactZones@data$Reference,function(x){
  temp<-SpatialPolygons((subset(impactZones,Reference==x)@polygons)) %>%
      areaPolygon() %>%
      as_data_frame() %>%
      mutate(Reference=x,
             area_km2=value/1e6,
             Control="Impact")
}) %>%
  bind_rows()

controlAreas <- lapply(controlZones@data$Reference,function(x){
  temp<-SpatialPolygons((subset(controlZones,Reference==x)@polygons)) %>%
      areaPolygon() %>%
      as_data_frame() %>%
      mutate(Reference=x,
             area_km2=value/1e6,
             Control="Control")
}) %>%
  bind_rows()

masterAreas <- bind_rows(impactAreas,controlAreas) %>%
  left_join(somaliPiracy@data %>%
              dplyr::select(Reference,AttackDate = Date))
```


## Plot locations of attacks

```{r plot}

leaflet(data = world) %>%
  # fitBounds(0, 60, 35, 72) %>%
  addPolygons(
    stroke = FALSE, fillOpacity = 1, smoothFactor = 1, color = "white"
  )  %>%
  addPolygons(
    data = eezSomalia,
    group = "eez",
    stroke=TRUE,
    weight=1,
    popup = ~paste0(
      "<b>EEZ: </b>", GeoName, "<br>")) %>%
  addMarkers(
    data = somaliPiracy,
    popup = ~paste0(
      "<b>Reference ID: </b>", Reference, "<br>",
       "<b>EEZ: </b>", GeoName, "<br>",
      "<b>Date: </b>", Date, "<br>",
      "<b>Aggressor: </b>", Aggressor, "<br>",
      "<b>Victim: </b>", Victim, "<br>",
      "<b>Description: </b>", gsub("NA","",paste0(Desc1,Desc2,Desc3,Desc4,Desc5,Desc6,sep=" ")), "<br>")) %>%
  addPolygons(data=impactZones,
              color="red",
              stroke=FALSE,
              fillOpacity = 0.5,
    popup = ~paste0(
      "<b>Reference ID: </b>", Reference, "<br>",
       "<b>EEZ: </b>", GeoName.1, "<br>",
      "<b>Zone type: </b>", Control, "<br>",
      "<b>Date: </b>", Date, "<br>",
      "<b>Aggressor: </b>", Aggressor, "<br>",
      "<b>Victim: </b>", Victim, "<br>",
      "<b>Description: </b>", gsub("NA","",paste0(Desc1,Desc2,Desc3,Desc4,Desc5,Desc6,sep=" ")), "<br>")) %>%
  addPolygons(data=controlZones,
              color="green",
              stroke=FALSE,
              fillOpacity = 0.5,
    popup = ~paste0(
      "<b>Reference ID: </b>", Reference, "<br>",
       "<b>EEZ: </b>", GeoName.2, "<br>",
      "<b>Zone type: </b>", Control, "<br>",
      "<b>Date: </b>", Date, "<br>",
      "<b>Aggressor: </b>", Aggressor, "<br>",
      "<b>Victim: </b>", Victim, "<br>",
      "<b>Description: </b>", gsub("NA","",paste0(Desc1,Desc2,Desc3,Desc4,Desc5,Desc6,sep=" ")), "<br>")) %>%
  fitBounds(xmin(extent(eezSomalia)), ymin(extent(eezSomalia)), xmax(extent(eezSomalia)), ymax(extent(eezSomalia)))
```

## Load GFW data

```{r GFWdata}
## Create bounding box around each zone
boxes <- lapply(controlZones@data$Reference,function(x){
  temp <- subset(controlZones,Reference == x)
  box <- bbox(temp) %>%
    as_data_frame
  
  box <- cbind(box[1,],box[2,])
  colnames(box) <- c("min_lon","max_lon", "min_lat","max_lat")
  box <- box %>%
    as_data_frame %>%
    mutate(Reference = x)
}) %>%
  bind_rows()

GFW_connection <- src_bigquery("ucsb-gfw", "fao_v2", max_pages = Inf)
fao_file <- "all_years_clean_effort"

#fishingEffort <- lapply(c("2016-209"), function(x){
fishingEffort <- lapply(somaliPiracy@data$Reference, function(x){
  print(x)
  tempBox <- boxes[which(boxes$Reference == x),]
  max_lat <- tempBox$max_lat
  min_lat <- tempBox$min_lat
  max_lon <- tempBox$max_lon
  min_lon <- tempBox$min_lon
  tbl(GFW_connection, from = fao_file) %>%
    dplyr::select(timestamp, lon, lat, hours, measure_new_score) %>%
    filter(
      lat < max_lat &
        lat > min_lat &
        lon < max_lon &
        lon > min_lon &
        measure_new_score > 0.5
    ) %>%
    collect(n=Inf) %>%
    mutate(Reference = tempBox$Reference)
})   %>%
  ## bind over all attacks
  bind_rows()

coordinates(fishingEffort) <- c("lon", "lat")
proj4string(fishingEffort) <- sp::proj4string(somaliPiracy)

daily_fishing_effort <- lapply(somaliPiracy@data$Reference,function(x){
  controlEffort <- raster::intersect(subset(fishingEffort,Reference == x),
                                     subset(controlZones,Reference == x)) %>%
    as_data_frame()
  
  impactEffort <- raster::intersect(subset(fishingEffort,Reference == x),
                                    subset(impactZones,Reference == x)) %>%
    as_data_frame()
  
  rbind(controlEffort,impactEffort)
}) %>%
  bind_rows() %>%
  mutate(Date = floor_date(timestamp, unit = "days")) %>%
  group_by(Reference,Control,Date) %>%
  summarize(Total_Hours = sum(hours)) %>%
  left_join(masterAreas,by=c("Reference","Control")) %>%
  mutate(hours_per_km2 = Total_Hours/area_km2)
```

## Statistics
Here we perform some statistics on the daily fishing effort data. These include:  
  
* Use a 30-day rolling mean for daily fishing effort [hours/km^2]
  
* Get predicted values from a regression that includes calendar effects, ENSO effect, and site-level effect  
  
* Detrend the data by subtracting this predicted regression from the observed values   
  
```{r stats,error=FALSE,message=FALSE,warning=FALSE}
## read in enso data
enso <- 
  read_table("http://www.esrl.noaa.gov/psd/gcos_wgsp/Timeseries/Data/nino34.long.anom.data",
             na = c("-99.99", "99.99"), skip = 1, n_max = year(Sys.time()) - 1870 + 1,
             col_names = c("year", 1:12)) %>%
  filter(year >= min(year(daily_fishing_effort$Date))) %>%
  filter(year <= max(year(daily_fishing_effort$Date))) %>%
  gather(month, enso, -year) %>%
  mutate(month = as.double(month))

## add column for 30 day rolling mean, join on enso data
daily_fishing_effort <- daily_fishing_effort %>%
  mutate(year = year(Date),
         month = month(Date)) %>%
  left_join(enso, by = c("year", "month"))

seas_pred <-
  daily_fishing_effort %>% 
  group_by(Reference,Control) %>%
  do(pred = predict(lm(hours_per_km2 ~ factor(month) + factor(wday(Date)), data = .))) %>%
  ungroup()
```

```{r resultsDailyEffort, fig.width=8.5, fig.height=6,error=FALSE,message=FALSE,warning=FALSE}
lapply(somaliPiracy@data$Reference,function(x){
  
  daily_fishing_effort_temp <- daily_fishing_effort %>%
    subset(Reference == x) 
  daily_fishing_effort_temp %>%
    ggplot(aes(x=as_date(Date),y=hours_per_km2,col=Control,area=Control)) +
    geom_line(lwd=0.8) +
    geom_vline(aes(xintercept = as.numeric(as_date(daily_fishing_effort$AttackDate[1]))),linetype=2,lwd=1.2,col="blue") +
  scale_x_date(limits = as.Date(c('2012-01-01','2016-12-31')))
})
```
